---
title: "Procedures and analysis plan - Quantifying the economic benefits of privately funded malaria control interventions in Maragra, Mozambique"
author: 
  - Joe Brew
  - Laia Cirera
  - Elisa Sicuri  
csl: health-economics-policy-and-law.csl
output:
  pdf_document:
    includes:
      in_header: style.sty
  html_document: default
highlight: zenburn
latex_engine: xelatex
bibliography: library.bib
---




```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE} 
#libraries
library(ggmap)
library(ggplot2)
library(grid)
library(dplyr)
library(knitr)
library(ggthemes)
library (rgdal)
library(RColorBrewer)
library(maptools)
opts_chunk$set(comment = NA, message = FALSE, echo = FALSE, results = 'hide', fig.height = 3.5, warning = FALSE)
```

```{r}
suppressWarnings(library(xtable))
options(xtable.comment = FALSE)
```

```{r}
# source('helpers.R')
```


\begin{changemargin}{3cm}{3cm} 

\begin{center}
\textbf{Summary}
\end{center}

\emph{The eradication of malaria could be accelerated by enlisting multiple non-governmental actors, including large private firms operating in endemic regions. Participation, however, relies on a clear and accurate quantification of the cost and impact of malaria control activities on economic output. This project proposes to use the case of the sugar mill of Maragra Açucar CA (southern Mozambique) to assess the impact of privately-managed indoor residual spraying on workers' economic output (in terms of both absenteeism and productivity). This document gives an overview of both the plan for data collection and analysis.}

\end{changemargin}

\newpage


\columnsbegin

# Background  


The burden of malaria is extremly high in Mozambique, even by regional standards [@World1999]. With a prevalence as high as 40%, malaria accounts for 29% of all deaths, and 42% of deaths among children under five [@dhs]. Nearly a quarter of maternal deaths are due to malaria [@Singh2014]. Along with HIV/AIDS [@Berg2014], malaria is one of the greatest threats to public health in southern Mozambique.  

In addition to malaria's impact on the health of its victims, the illness also has major economic consequences for the ill. Children who survive malaria face hurdles which can have life-long economic repurcussions, particularly those related to intellectual development (such as cerebral malaria) [@Idro2010] and general growth anemia [@Mabunda2008]. Their families also pay economically - 32-34% of households incur malaria-related costs which rise to the level of "catastrohpic" per the World Health Organization's standards (ie, 10% of household income or 40% of non-food income) [@CastilloRiquelme2008]. Though the burden of malaria is decreasing [@Murray2014], the costs of the disease at the individual level remain enormous, given that the disease affects primarily those with low socioeconomic status.  

The economic effects of malaria are not only absorped but its direct victims, but also by the economy as a whole. Malaria control has been found to be associated with population-level economic growth in multiple studies [@Barofsky2015]. By eliminating early-life blocks on the development of a population's human capital, the returns on a reduction in malaria's burden are long-term and exponential.  

From a public health perspective, the case for the need to control and eventually eradicate malaria is strong and has been made clear in multiple studies across time and geography. However, the role which private firms which operate in malaria endemic regions can be expected to play is less obvious, given the current lack of compelling evidence regarding the return on investment in short- and medium-terms for privately-funded malaria control activities. To the extent that many firms already carry out some form of "in-house" malaria control, analyzing those firms' data offers the unique opportunity to assess whether the benefits (in purely economic terms) of those activities outweigh the costs, or vice-versa.

Findings of the few studies performed are discordant, pointing either to higher earnings and higher literacy rates in adults born after malaria campaigns or to non significant results. The value of these studies to inform policy making processes is, thus, questionable.     

# Study Rationale

In recent years, much of the discourse regarding malaria in endemic regions has shifted from control to eradication [@Tanner2015]. This shift will require changes not only in the implementation of public health interventions, but also the recruitment and sustenance of multiple and novel fronts in the fight against malaria. One potential new front and source of funding would be private firms, especially given the astounding growth of foreign direct investment in Mozambique over the last two decades.

Though private, foreign-owned firms are a potential source of funding for malaria control and eradication, it is not reasonable to expect significant participation at the population level without a clear demonstration of the value proposition from a private perspective. Though the non-tangible benefits of "corporate social responsibility" (good publicity, etc.) are certainly appealing to private firms, investment would likely increase significantly if the costs and benefits of malaria control from a purely economic perspective were quantified. On the one hand, if it can be compellingly demonstrated that privately-funded malaria control interventions offer a significant return on investment, this evidence could entail an organic/spontaneous investment increase across the country. On the other hand, if it were found that private foreign investment is not cost-effective from the perspective of private foreign firms, then this may spur donors and the public sector to better coordinate, scale, and work with firms to fill the gaps.


Evaluation analyses of the effect of malaria control on indicators of economic growth deriving from an actual program may bring a unique contribution to the literature. Most likely, it would clarify if malaria control interventions have a significant impact on labour productivity and would provide important information for future health policy decisions. 

As a consequence of the malaria control activities undertaken, in conjunction with the Ministries of Health of the countries in which the company operates Maragra Açucar in Mozambique as well as all the other sugar Illovo factories operating in different countries, is likely to benefit not only the health of its workers and of their families but also their capacity to work and to contribute to the economic growth and development of the area. Measuring this potential impact now is of paramount importance as this may serve as baseline information for the evaluation of a future elimination programme in the area. 


Ultimately, the justification for this study is opportunity. The presence in Mozambique of large, foreign firms with sophisticated systems for inventory, employee activities, and (in some cases) even on-site clinics means that, for the first time, there exist data which can shed light on the feasibility and scalability scaling up non-traditional fronts in the fight against malaria.


# Objectives  

## Primary objective  

The primary objective is to identify and measure the impact of malaria interventions on selected economic indicators.


## Specific objectives

In addition, and in complentarity to the primary objective, secondary objectives include:  

- The quantification of the direct economic costs and benefits of malaria interventions carried out by private firms, using the case of Maragra Açucar CA.
- The quantification of worker absenteeism over time, and in causal association with malaria control activities.
- The quantification of worker _productivity_ over time, and in causal association with malaria control activities.
- The quantification of worker _illness_ over time, specifically in regards to those diseases directly related to vector control, and the corresponding estimation of the direct treatment costs.
- The estimation of costs associated with privately-funded malaria control activities.
- The overall and intervention-specific return on investment for privately-funded malaria control activities.

# Methods  


## General  

We plan to identify the relationship between malaria and workers’ absenteeism from Maragra Açucar (and from the other factories) by comparing information possibly routinely recorded by the company before and after activities of malaria control (such as fumigation or bed net distribution or other) are carried out. A further comparison could be undertaken between absenteeism of workers living in areas with different levels of malaria incidence or simply with different timing of the implementation of malaria control activities. Obtaining information on the reason of workers’ absenteeism constitutes additional value in this analysis. In order to assess the actual impact of malaria on absenteeism, it is essential to control for confounding factors that may also affect absenteeism, such as workers’ age, socioeconomic status, living place, and many others.

### The data 

In the spring and summer of 2016, under the auspices of the MALTEM (Mozambican Alliance Towards the Elimination of Malaria), and in cooperation with the CISM (Centro de Investigaçao em Saude de Manhiça) and Maragra Açucar CA, data will be collected. Individual information on absenteeism (possibly with the reason of absenteeism), age, gender, job position, workers’ place of residence, educational level, information on whether the worker is a migrant and if yes, when and how long the worker lives and distance to job are all of great importance for this study. In the best case scenario, a complete panel data (e.g., monthly individual data) would allow to discern how malaria control activities are translated into the outcomes of interest, in time.

For the above, the following specific data will be required:  


- All vector-control activities carried out by Maragra (indoor residual spraying and bednets distribution), specifically in relation to time and location.
- The costs (both direct and indirect) associated with the above activities.
- Worker absenteeism data.
- Clinic data pertaining to worker illnesses / visits.
- Worker sociodemographic and residential data.
- Worker performance/output data (if available).
- Worker wages and indirect costs (if available).

Additionally, to account for confounding factors in terms of both population characteristics and interventions, both census and MALTEM activities data will be utilized.

Finally, to account for outsides factors, precipitation and temperature data will be used for the purposes of modeling and adjusting for seasonlity.

### Study area  

The study area pertains to the sugar plantation and factory of Maragra.

### Study design  

This study is purely retrospective, using existing data sources, with no intervention. It relies on the "natural experiment" of differential applications of a treatment (IRS) by time and geography in relation to economic outcomes (absenteeism and productivity). It has both a descriptive (correlational) component, as well an analytical component (a retrospective cohort study).

### Study population  

The study population consists of all workers of Maragra Açucar CA from the period of 2010 through 2015.

### Timing and duration of the study  

Data collection will take place in May-June of 2016. Analysis will be carried out in July-August of 2016. Results and manuscript writing can be expected to be finished by the end of August 2016. Internal vetting of manuscript and submission for publication will be finished by the end of September, 2016.  

### Data collection and management  

All data are retrospective and already in existence. In this sense, data are to be "compiled" rather than collected.  Compiled data will be stored on an encrypted drive to which only the PI (Elisa Sicuri) and researcher (Joe Brew) will have access. 

### Data analysis  

Two analytical approaches are envisioned for the impact evaluation and cost quantification of Maragra's malaria control initiatives on economic outcomes, respectively:  

1. A "difference in differences" approach to assess the impact of IRS in terms of the aggregated before/after blocks, accounting for omitted variable bias. 

```{r}
a <- seq(10, 7, length = 100)
b <- c(seq(9, 7.5, length = 50), (seq(6, 4.5, length = 50)))

plot(x = 1:100,
     y = a,
     type = 'l',
     ylim = c(0,11),
     xlab = 'Time',
     ylab = 'Absenteeism')
lines(x = 1:100,
      y = b,
      col = 'red')
text(x = 12, 
     y = 10.5,
     label = 'Control group',
     cex = 1)
text(x = 15, 
     y = 8,
     label = 'Treatment group',
     col = 'Red',
     cex = 1)
text(x = 50,
     y = 6.5,
     label = 'Effect of treatment',
     cex = 0.8)
title(main = 'Example of difference in differences methodology')

```

2. Logistic regression to assess the likelihood of worker absence as a function of time since indoor residual spraying, adjusting for confouding variables (weather/seasonality, sociodemogrpahic characteristics, worker type, etc.), thereby assessing IRS' waning impact over time.  

```{r}
vals <- c(-0.3, -0.13, 0.1, 0.4, 0.2, -0.12)
bp <- barplot(vals,
              names.arg = gsub(' ', '\n', c('IRS', 'Bednet', 'Male', 'Rainy season', 'Agricultural worker', 'Non-resident')),
              cex.names = 0.7,
              border = NA,
              xlab='Characteristic',
              ylab = 'Effect on likelihood of absenteeism',
              cex.lab = 1.2)
title(main = 'Example of logistic regression methodology')
```

Following these two analyses, the extension to the costing portion will be relatively simple, requiring only basic arithmetic and counterfactual scenario generations.

### Ethics clearance  

This project falls under the auspices of the already-approved protocol of the MALTEM economic evaluation.  

### Confidentiality  

All data will be stored on encryped disks, and identifiable data will only be accessible to the PI and main researcher.

# Details on and qualifications of research team  

Details on members of the research team are available at [www.economicsofmalaria.com](www.economicsofmalaria.com).   




```{r, eval = FALSE}
library(raster)
# Get map of mozambique
moz <- getData('GADM', country = 'MOZ', level = 3)
# Get map of manhica district
man <- moz[moz$NAME_2 == 'Manhiça',]
# Make some random points within man
x <- bbox(man)
xs <- sample(seq(x[1], x[3], length = 100000), 10000)
ys <- sample(seq(x[2], x[4], length = 100000), 10000)
random <- data.frame(x = xs,
                     y = ys)
coordinates(random) <- ~x+y
# Widdle down those that aren't in man
random <- SpatialPointsDataFrame(coords = coordinates(random),
                                 data = data.frame(id = seq(1:10000)))
proj4string(random) <- proj4string(man)
random <- random[!is.na(over(random, polygons(man))),]
# Give values to random's dataframe
random$long <- coordinates(random)[,1]
random$lat <- coordinates(random)[,2]

# Get the specific localidade
random@data$localidade <- man@data$NAME_3[over(random, polygons(man))]

# First, assign a few hotspots
hotspots <- random[sample(1:nrow(random), 10, replace = TRUE),]
# Define whether a location is within x distance from a hotspot
# in_hotspot <- rep(FALSE, nrow(random))
hotspot_mat <- matrix(rep(NA), ncol = nrow(hotspots), nrow = nrow(random))
for (i in 1:nrow(hotspots)){
  temp <- spDistsN1(pts = random, pt = hotspots[i,])  
  hotspot_mat[,i] <- temp
  # in_hotspot[temp < 0.1] <- TRUE
}
# Add mean distance from hotspots to random
random$mean_distance_from_hotspot <- apply(hotspot_mat, 1, mean)
# Conditionally give more weight to those in hotspots
random$coverage <- random@data$mean_distance_from_hotspot +
  rnorm(n = nrow(random), mean = 0, sd = 0.05) + 0.2
random$value <- as.numeric(random$coverage >= 0.5)
# Add some arbitrary household sizes
random$hh_size <- sample(1:15, nrow(random), replace = TRUE)

# Get mean values from random into man
man@data <- man@data %>%
  mutate(localidade = NAME_3) %>%
  left_join(random@data %>%
  group_by(localidade) %>%
  summarise(coverage = mean(coverage)))

# Fortify objects
man_fort <- fortify(man, region = 'NAME_3') %>%
  mutate(localidade = id) %>%
  left_join(man@data %>%
              dplyr::select(localidade, coverage))



```

```{r, eval = FALSE}
# RAW POINTS
# 
# Make map
ggplot(man_fort,
       aes(x = long, y = lat)) +
  geom_polygon(aes(group = id), fill = 'grey') +
  geom_point(data = random@data,
             aes(x = long, y = lat, col = coverage * 10), size = 0.35) +
  coord_map() +
  geom_polygon(aes(group = id), fill = NA, color = 'black') +
  scale_colour_gradient(low = 'darkorange',
                      # guide = guide_legend(reverse=TRUE), 
                      name = 'Meticais') +
  xlab('Longitude') +
  ylab('Latitude') +
  ggtitle('Example of results: perceived monetary value of IRR') +
  theme_gg
```

```{r, eval = FALSE}
# Localidade-scores

ggplot(man_fort,
       aes(x = long, y = lat)) +
  geom_polygon(aes(group = id, fill = coverage * 10)) +
  coord_map() +
  geom_polygon(aes(group = id), fill = NA, color = 'black') +
  scale_fill_gradient(low = 'darkorange',
                      # guide = guide_legend(reverse=TRUE), 
                      name = 'Vaccine aversion index') +
  xlab('Longitude') +
  ylab('Latitude') +
  ggtitle('Example of results: vaccine receptivity') +
  theme_gg
```


\columnsend


# References  



