---
title: "Procedures and analysis plan - Quantifying the economic benefits of privately funded malaria control interventions in Maragra, Mozambique"
author: 
  - Joe Brew
  - Laia Cirera
  - Elisa Sicuri  
csl: health-economics-policy-and-law.csl
output:
  pdf_document:
    includes:
      in_header: style.sty
  html_document: default
highlight: zenburn
latex_engine: xelatex
bibliography: library.bib
---




```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE} 
#libraries
library(ggmap)
library(ggplot2)
library(grid)
library(dplyr)
library(knitr)
library(ggthemes)
library (rgdal)
library(RColorBrewer)
library(maptools)
opts_chunk$set(comment = NA, message = FALSE, echo = FALSE, results = 'hide', fig.height = 6, warning = FALSE, fig.width = 6)
```

```{r}
suppressWarnings(library(xtable))
options(xtable.comment = FALSE)
```

```{r}
source('helpers.R')
```


\begin{changemargin}{3cm}{3cm} 

\begin{center}
\textbf{Summary}
\end{center}

\emph{The eradication of malaria could be accelerated by enlisting multiple non-governmental actors, including large private firms operating in endemic regions. Participation, however, relies on a clear and accurate quantification of the cost and impact of malaria control activities on economic output. This project proposes to use the case of the sugar mill of Maragra Açucar CA (southern Mozambique) to assess the impact of privately-managed indoor residual spraying on workers' economic output (in terms of both absenteeism and productivity). This document gives an overview of both the plan for data collection and analysis.}

\end{changemargin}

\newpage


\columnsbegin

# Background  


The burden of malaria is extremly high in Mozambique, even by regional standards [@World1999]. With a prevalence as high as 40%, malaria accounts for 29% of all deaths, and 42% of deaths among children under five [@dhs]. Nearly a quarter of maternal deaths are due to malaria [@Singh2014]. Along with HIV/AIDS [@Berg2014], malaria is one of the greatest threats to public health in southern Mozambique.



In addition to malaria's impact on the health of its victims, the illness also has major economic consequences for the ill. Children who survive malaria face hurdles which can have life-long economic repurcussions, particularly those related to intellectual development (such as cerebral malaria) [@Idro2010] and general growth anemia [@Mabunda2008]. Their families also pay economically - 32-34% of households incur malaria-related costs which rise to the level of "catastrohpic" per the World Health Organization's standards (ie, 10% of household income or 40% of non-food income) [@CastilloRiquelme2008]. Though the burden of malaria is decreasing [@Murray2014], the costs of the disease at the individual level remain enormous, given that the disease affects primarily those with low socioeconomic status.

The economic effects of malaria are not only absorped but its direct victims, but also by the economy as a whole.  ....

- *Malaria is bad for firms*  


# Study Rationale

*Control of malaria is a potential win/win for business/government*  

*For firms to step up, compelling evidence quantifying their potential benefits is needed*  


# Objectives  

## Primary objective  

Quantify the direct economic benefits of malaria intervention

## Specific objectives



# Methods

## General  

### Study area  

### Study design  

### Study population  

### Timing and duration of the study  

### Data collection and management  

### Data analysis  

### Ethics clearance  

### Confidentiality  

## Methodology specifications by objective  


# Details on and qualifications of research team  

Details on members of the research team are available at [www.economicsofmalaria.com](www.economicsofmalaria.com).   


# References  

# Appendix  


```{r}
library(raster)
# Get map of mozambique
moz <- getData('GADM', country = 'MOZ', level = 3)
# Get map of manhica district
man <- moz[moz$NAME_2 == 'Manhiça',]
# Make some random points within man
x <- bbox(man)
xs <- sample(seq(x[1], x[3], length = 100000), 10000)
ys <- sample(seq(x[2], x[4], length = 100000), 10000)
random <- data.frame(x = xs,
                     y = ys)
coordinates(random) <- ~x+y
# Widdle down those that aren't in man
random <- SpatialPointsDataFrame(coords = coordinates(random),
                                 data = data.frame(id = seq(1:10000)))
proj4string(random) <- proj4string(man)
random <- random[!is.na(over(random, polygons(man))),]
# Give values to random's dataframe
random$long <- coordinates(random)[,1]
random$lat <- coordinates(random)[,2]

# Get the specific localidade
random@data$localidade <- man@data$NAME_3[over(random, polygons(man))]

# First, assign a few hotspots
hotspots <- random[sample(1:nrow(random), 10, replace = TRUE),]
# Define whether a location is within x distance from a hotspot
# in_hotspot <- rep(FALSE, nrow(random))
hotspot_mat <- matrix(rep(NA), ncol = nrow(hotspots), nrow = nrow(random))
for (i in 1:nrow(hotspots)){
  temp <- spDistsN1(pts = random, pt = hotspots[i,])  
  hotspot_mat[,i] <- temp
  # in_hotspot[temp < 0.1] <- TRUE
}
# Add mean distance from hotspots to random
random$mean_distance_from_hotspot <- apply(hotspot_mat, 1, mean)
# Conditionally give more weight to those in hotspots
random$coverage <- random@data$mean_distance_from_hotspot +
  rnorm(n = nrow(random), mean = 0, sd = 0.05) + 0.2
random$value <- as.numeric(random$coverage >= 0.5)
# Add some arbitrary household sizes
random$hh_size <- sample(1:15, nrow(random), replace = TRUE)

# Get mean values from random into man
man@data <- man@data %>%
  mutate(localidade = NAME_3) %>%
  left_join(random@data %>%
  group_by(localidade) %>%
  summarise(coverage = mean(coverage)))

# Fortify objects
man_fort <- fortify(man, region = 'NAME_3') %>%
  mutate(localidade = id) %>%
  left_join(man@data %>%
              dplyr::select(localidade, coverage))



```

```{r, eval = FALSE}
# RAW POINTS
# 
# Make map
ggplot(man_fort,
       aes(x = long, y = lat)) +
  geom_polygon(aes(group = id), fill = 'grey') +
  geom_point(data = random@data,
             aes(x = long, y = lat, col = coverage * 10), size = 0.35) +
  coord_map() +
  geom_polygon(aes(group = id), fill = NA, color = 'black') +
  scale_colour_gradient(low = 'darkorange',
                      # guide = guide_legend(reverse=TRUE), 
                      name = 'Meticais') +
  xlab('Longitude') +
  ylab('Latitude') +
  ggtitle('Example of results: perceived monetary value of IRR') +
  theme_gg
```

```{r}
# Localidade-scores

ggplot(man_fort,
       aes(x = long, y = lat)) +
  geom_polygon(aes(group = id, fill = coverage * 10)) +
  coord_map() +
  geom_polygon(aes(group = id), fill = NA, color = 'black') +
  scale_fill_gradient(low = 'darkorange',
                      # guide = guide_legend(reverse=TRUE), 
                      name = 'Vaccine aversion index') +
  xlab('Longitude') +
  ylab('Latitude') +
  ggtitle('Example of results: vaccine receptivity') +
  theme_gg
```

\columnsend
